---
title: "Long run database test"
author: "Matthew Kuperus Heun"
date: '`r format(Sys.time(), "%d %B %Y")`'
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```


# Setup

Load required packages.

```{r}
library(targets)
# remove the `_targets_r` directory previously written by non-interactive runs of the report.
tar_unscript()
```


# Globals

Define global options and functions.

```{targets example-globals, tar_globals = TRUE}
options(tidyverse.quiet = TRUE)
tar_option_set(packages = c("dplyr", "ggplot2", "readxl", "tidyr", "matsindf", "Recca"))

read_data <- function(.path, sheet) {
  .path |> 
    readxl::read_excel(sheet = sheet) 
}
```

# Targets

Load the raw Iron and steel data from Ricardo.

```{targets iron_steel, tar_simple = TRUE}
read_data("data/Iron_Steel_data_US_2011_2001.xlsx", sheet = "Iron and steel")
```

Process the data into a `matsindf` data frame, 
a.k.a., a magic spreadsheet.
First step is to make S_units vectors.

```{targets s_units, tar_simple = TRUE} 
iron_steel |> 
  dplyr::mutate(
    # Create a "Product" as expected when creating the S_units matrix.
    "{IEATools::iea_cols$product}" := dplyr::case_when(
      .data[["Matrix"]] %in% c("R", "V") ~ .data[["To"]], 
      .data[["Matrix"]] %in% c("U", "Y") ~ .data[["From"]], 
      TRUE ~ NA_character_
    ), 
    # Create a "Flow" as expected when creating the S_units matrix.
    "{IEATools::iea_cols$flow}" := dplyr::case_when(
      .data[["Matrix"]] %in% c("R", "V") ~ .data[["From"]], 
      .data[["Matrix"]] %in% c("U", "Y") ~ .data[["To"]], 
      TRUE ~ NA_character_
    ), 
    # Eliminate unneeded columns.
    From = NULL, 
    To = NULL, 
    Matrix = NULL
  ) |> 
  # Now make the S_units vectors
  IEATools::extract_S_units_from_tidy(matnames = "Matrix", e_dot = "Value")
```
Then create RUVY matrices in a data frame.

```{targets ruvy_df, tar_simple = TRUE}
iron_steel |> 
  dplyr::group_by(Year, Country, Energy.type, Matrix) |> 
  matsindf::collapse_to_matrices(matnames = "Matrix", 
                                 rownames = "From", 
                                 colnames = "To",
                                 matvals = "Value",
                                 matrix_class = "matrix") |> 
  tidyr::pivot_wider(names_from = "Matrix", values_from = "Value") |> 
  dplyr::mutate(
    # Set row and column types
    R = R |> matsbyname::setrowtype("Industry") |> matsbyname::setcoltype("Product"), 
    U = U |> matsbyname::setrowtype("Product") |> matsbyname::setcoltype("Industry"), 
    V = V |> matsbyname::setrowtype("Industry") |> matsbyname::setcoltype("Product"), 
    Y = Y |> matsbyname::setrowtype("Product") |> matsbyname::setcoltype("Industry"), 
    # Create other expected matrices
    # The U_feed matrix is same as U
    U_feed = U, 
    # Create a U_EIOU matrix full of zeros but with same rows and columns as U
    U_EIOU = matsbyname::hadamardproduct_byname(U, 0),
    # Create an r_EIOU matrix, again full of zeroes
    r_EIOU = U_EIOU
  ) |> 
  # Join with the S_units matrices
  dplyr::left_join(s_units, by = c("Year", "Country", "Energy.type"))
```



# Pipeline

If you ran all the `{targets}` chunks in non-interactive mode, then your R scripts are set up to run the pipeline.

```{r}
tar_make()
```

The `targets` dependency graph shows the steps of the calculation pipeline.

```{r}
tar_visnetwork()
```


# Output

Retrieve results from the `_targets/` data store using `tar_read()`.

```{r, message = FALSE}
df <- tar_read(ruvy_df)
df
```

We can dump the matrices to an Excel file for further interrogation.

```{r}
df |> 
  Recca::write_ecc_to_excel("~/Desktop/IronAndSteelRUVYMatrices.xlsx", 
                            overwrite_file = TRUE)
```

Look at some of the matrices here.

The resource matrix (**R**) contains exogenous inputs to the energy conversion chain.
Imports can also be included here for a single country.
The next bit of code grabs the **R** matrix in the first row (2001).

```{r}
df$R[[1]]
```

The use matrix (**U**) contains all uses of energy by transformation
industries in the economy.
The next bit of code grabs the **U** matrix in the second row (2011).

```{r}
df$U[[2]]
```

The make matrix (**V**) contains all uses of energy by transformation
industries in the economy.
The next bit of code grabs the **V** matrix in the first row (2001).

```{r}
df$V[[1]]
```

The final demand matrix (**Y**) shows sectoral uses of energy.
If we had more than the "Iron and steel" sector,
the **Y** matrix would contain additional columns.
The next bit of code grabs the **Y** matrix in the second row (2011).

```{r}
df$Y[[2]]
```


We can check whether the matrices exhibit energy balance.

```{r}
df |> 
  Recca::verify_SUT_energy_balance() |> 
  dplyr::select(Country, Year, .SUT_energy_balance)
```

We can visualize with Sankey diagrams. 
This code makes a Sankey diagram from each row of the data frame.

```{r}
sankeys <- df |> 
  Recca::make_sankey()
sankeys
```

Here is the Sankey diagram for the first row (2001).

```{r}
sankeys |> 
  dplyr::filter(Year == 2001) |> 
  magrittr::extract2("Sankey") |> 
  magrittr::extract2(1)
```

Here is the Sankey diagram for the second row (2011).

```{r}
sankeys |> 
  dplyr::filter(Year == 2011) |> 
  magrittr::extract2("Sankey") |> 
  magrittr::extract2(1)
```



# Analysis

We can run all types of analyses after the data are in the RUVY format.
For example, this code calculates the efficiency of all Industries.

```{r}
eta_i <- df |>
  Recca::calc_eta_i()
```

The **eta_i** vectors contain efficiencies for each processing stage (Industry) 
in the energy conversion chain.
Here are the efficiencies for 2001.

```{r}
eta_i$eta_i[[1]]
```

And here are the efficiencies for 2011.

```{r}
eta_i$eta_i[[2]]
```

Note that the code to calculate efficiencies applies to the `df` data frame.
Thus, the same code works for any number of rows in the data frame.
I.e., the same code works for any number of countries and years.

Furthermore, the code to calculate efficiencies applies to the **U** and **V** matrices,
regardless of their size. 
So, the same code works if we have more energy carriers or more 
energy transformation industries.


# Questions and comments

(0) Things are looking really good!
(1) The efficiency of "Electronics" is `INF`. And the Sankey diagrams
    show energy created by "Electronics" but no energy consumed by "Electronics".
    That's probably not what we want.
(2) "Commercial_indoor_electric_lamps", "Electric_arc_furnaces", and 
    "Room_air_conditioners" have underscore characters ("_") in their names.
    The Short-run database does not have underscore characters in names.
    So maybe remove those characters?
(3) Check the Excel file.  You'll find nicely formatted matrices in Excel format.
    Those may be helpful for debugging.
(4) Note that I added **U_EIOU** and **U_feed** matrices. 
    **U_EIOU** is zero for now, but we can add energy industry own use later.
    Because **U_EIOU** is the **0** matrix, **r_EIOU** is also **0** (for now).
(5) I changed the name of the tab to be "Iron and steel", assuming 
    you will probably have other tabs in the future.
(6) As I advised in an earlier call, I would not divorce the file read by `R`
    from the raw data you used to create the data read by `R`.

    
