---
title: "Long run database test"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```


# Setup

Load required packages.

```{r}
library(targets)
# remove the `_targets_r` directory previously written by non-interactive runs of the report.
tar_unscript()
```


# Globals

Define global options and functions.

```{targets example-globals, tar_globals = TRUE}
options(tidyverse.quiet = TRUE)
tar_option_set(packages = c("dplyr", "ggplot2", "readxl", "tidyr", "matsindf"))

read_data <- function(.path, sheet) {
  .path |> 
    readxl::read_excel(sheet = sheet) 
}
```

# Targets

Load the raw Iron and steel data from Ricardo.

```{targets iron_steel, tar_simple = TRUE}
read_data("data/Iron_Steel_data_US_2011_2001.xlsx", sheet = "Iron and steel")
```

Process the data into a `matsindf` data frame, 
a.k.a., a magic spreadsheet.

```{targets ruvy_df, tar_simple = TRUE}
iron_steel |> 
  dplyr::group_by(Year, Country, Energy.type, Matrix) |> 
  matsindf::collapse_to_matrices(matnames = "Matrix", 
                                 rownames = "From", 
                                 colnames = "To",
                                 matvals = "Value") |> 
  tidyr::pivot_wider(names_from = "Matrix", values_from = "Value") |> 
  # Set row and column types
  dplyr::mutate(
    R = R |> matsbyname::setrowtype("Industry") |> matsbyname::setcoltype("Product"), 
    U = U |> matsbyname::setrowtype("Product") |> matsbyname::setcoltype("Industry"), 
    V = V |> matsbyname::setrowtype("Industry") |> matsbyname::setcoltype("Product"), 
    Y = Y |> matsbyname::setrowtype("Product") |> matsbyname::setcoltype("Industry")
  )
```

# Pipeline

If you ran all the `{targets}` chunks in non-interactive mode, then your R scripts are set up to run the pipeline.

```{r}
tar_make()
```

# Output

Retrieve results from the `_targets/` data store using `tar_read()` or `tar_load()`.

```{r, message = FALSE}
df <- tar_read(ruvy_df)
```

Look at some of the matrices.

The resource matrix (**R**) contains exogenous inputs to the energy conversion chain.
Imports can als be included here.
```{r}
df$R[[1]]
```

The use matrix (**U**) contains all uses of energy by transformation
industries in the economy.


The final demand matrix (**Y**) shows sectoral uses of energy.
If we had more than the "Iron and steel" sector,
the **Y** matrix would contain additional columns.

```{r}
df$Y[[2]]
```

The `targets` dependency graph shows the steps of the calculation pipeline.

```{r}
tar_visnetwork()
```

At this point, you can go back and run `{targets}` chunks in interactive mode without interfering with the code or data of the non-interactive pipeline.
